### 2020-04-09 ks/ks.Rmd
Can create simulated KS figure

---
title: "Kolmogorov-Smirnov"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(testthat)
library(ggplot2)
```

Human samples:

```{r}
set.seed(314)
# n: sample size, equals the number of HLA haplotypes
n <- 13
# f denotes 'fraction', a value between 0.0 and 1.0
f_human <- rnorm(n = n, mean = 0.053, sd = 0.01)
f_covid <- rnorm(n = n, mean = runif(n = 1, min = 0.03, max = 0.1), sd = 0.01)
expect_true(all(f_human > 0.0))
expect_true(all(f_human < 1.0))
expect_true(all(f_covid > 0.0))
expect_true(all(f_covid < 1.0))
```

Plot:

```{r}
df <- tibble::tibble(
  proteome = rep(c("human", "covid"), each = n),
  haplotype = rep(LETTERS[1:n], time = 2),
  f = c(f_human, f_covid)
)
df$proteome <- as.factor(df$proteome)
df$haplotype <- as.factor(df$haplotype)
ggplot(df, aes(x = haplotype, y = f, fill = proteome)) + 
  scale_fill_manual(values = c("human" = "#ffffff", "covid" = "#cccccc")) +
  geom_col(position = position_dodge(), color = "#000000") + xlab("HLA haplotype") + 
  ylab("% epitopes overlapping \nwith transmembrane helix") +
  geom_hline(yintercept = mean(f_human), lty = "dashed") +
  geom_hline(yintercept = mean(f_covid), lty = "dotted") +
  labs(
    title = "SIMULATED DATA!",
    caption = paste0(
      "Dashed line: average in humans: ", formatC(mean(f_human), digits = 2),". ",
      "Dotted line: average in COVID-19: ", formatC(mean(f_covid), digits = 2),"."  
    )
  ) + ggsave("fig_1.png", width = 7, height = 7)
```

Test: