### 2020-04-09 ks/ks.Rmd
Render KS to supmat

---
title: "Kolmogorov-Smirnov"
author: Richel J.C. Bilderbeek
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(testthat)
library(tidyr)
library(ggplot2)
```

Human samples:

```{r}
# n: sample size, equals the number of HLA haplotypes
n <- 13
# f denotes 'fraction', a value between 0.0 and 1.0
f_human <- rnorm(n = n, mean = 0.053, sd = 0.01)
f_covid <- rnorm(n = n, mean = runif(n = 1, min = 0.03, max = 0.1), sd = 0.01)
expect_true(all(f_human > 0.0))
expect_true(all(f_human < 1.0))
expect_true(all(f_covid > 0.0))
expect_true(all(f_covid < 1.0))
```

Test:

```{r}
t <- ks.test(x = f_human, y = f_covid)
print(t)
```

Stats:

```{r}
alpha <- 0.05
if (t$p.value <= alpha) {
  verdict <- "Reject that the two samples are taken from a same distribution"
} else {
  verdict <- "Accept that the two samples are taken from a same distribution"
}
print(verdict)
```

Create tidy table:

```{r}
df <- tibble::tibble(
  proteome = rep(c("human", "covid"), each = n),
  haplotype = rep(LETTERS[1:n], time = 2),
  f = c(f_human, f_covid)
)
df$proteome <- as.factor(df$proteome)
df$haplotype <- as.factor(df$haplotype)
```

## Stats output

```{r}
names(t)
df_t_wide <- data.frame(
  statistic = t$statistic,
  p_value = t$p.value,
  alternative = t$alternative,
  method = t$method,
  data_name = t$data.name,
  alpha = alpha,
  verdict = verdict
)
df_t <- t(df_t_wide)
values <- df_t[,1]
param_names <- rownames(df_t)
df_t <- data.frame(
  parameter = param_names,
  value = values
)

knitr::kable(df_t, row.names = FALSE)
```

```{r}
write.csv(df_t, "ks.csv", row.names = FALSE)
```

## Plot

```{r}
ggplot(df, aes(x = haplotype, y = f, fill = proteome)) + 
  scale_fill_manual(values = c("human" = "#ffffff", "covid" = "#cccccc")) +
  geom_col(position = position_dodge(), color = "#000000") + xlab("HLA haplotype") + 
  ylab("Epitopes overlapping \nwith transmembrane helix") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 3)) +
  geom_hline(yintercept = mean(f_human), lty = "dashed") +
  geom_hline(yintercept = mean(f_covid), lty = "dotted") +
  labs(
    caption = paste0(
      "Dashed line: average in humans: ", formatC(100.0 * mean(f_human), digits = 2),"%. ",
      "Dotted line: average in COVID-19: ", formatC(100.0 * mean(f_covid), digits = 2),"%."  
    )
  ) + ggsave("fig_1.png", width = 7, height = 7)
```

## Table

Create wide (untidy) table:

```{r}
df_wide <- spread(df, proteome, f)
df_wide$covid <- formatC(100.0 * df_wide$covid, digits = 3)
df_wide$human <- formatC(100.0 * df_wide$human, digits = 3)
```

Show table:

```{r}
knitr::kable(df_wide, caption = "Percentage of MHC-I epitopes overlapping with transmembrane helix")
```

Save table:

```{r}
write.csv(df_wide, "table_1.csv", row.names = FALSE)
```