### 2020-04-20 hydrophobicity/hydrophobicity.Rmd
Start exploring hydrophobicity

---
title: "Hydrophobicity"
author: "Richel J.C. Bilderbeek"
date: "4/20/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```