### 2020-05-19 covid_dataset_strength/covid_dataset_strength.Rmd
First start of determining the strength of the dataset

---
title: "COVID-19 dataset strength"
author: Richel J.C. Bilderbeek
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(testthat)
library(pureseqtmr)
```

Must be installed:

```{r}
expect_true(is_pureseqtm_installed())
```

```{r cache=TRUE}
fasta_filename <- "covid.fasta"
expect_true(file.exists(fasta_filename))
topology <- pureseqtmr::predict_proteome_topology(fasta_filename)
```

```{r}
n_non_tmh_aas <- sum(stringr::str_count(topology$topology, "0"))
message("number of non-TMH amino acids, n_non_tmh_aas: ", n_non_tmh_aas)
```

```{r}
n_tmh_aas <- sum(stringr::str_count(topology$topology, "1"))
message("number of TMH amino acids, n_tmh_aas: ", n_tmh_aas)
```

```{r}
n_aas <- n_non_tmh_aas + n_tmh_aas
message("proteome size, number of amino acids, n_aas: ", n_aas)
```

```{r}
perc_aas <- n_tmh_aas / (n_tmh_aas + n_non_tmh_aas)
message("percentage of proteome that is TMH, perc_aas: ", perc_aas)
```

```{r}
n_tmhs <- sum(tally_tmhs(topology)$n_tmhs) 
message("number of TMHs, n_tmhs: ", n_tmhs)
```