### 2020-05-20 compare_proteomes/compare_proteomes.Rmd
Copy proteomes from 'downloads' folder

---
title: "COVID-19 dataset strength"
author: Richel J.C. Bilderbeek
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(testthat)
library(pureseqtmr)
```

Must be installed:

```{r}
expect_true(is_pureseqtm_installed())
```

```{r}
show_properties <- function(fasta_filename) {
  expect_true(file.exists(fasta_filename))
  topology <- pureseqtmr::predict_proteome_topology(fasta_filename)
  n_proteins <- nrow(topology)
  n_non_tmh_aas <- sum(stringr::str_count(topology$topology, "0"))
  n_tmh_aas <- sum(stringr::str_count(topology$topology, "1"))
  n_aas <- n_non_tmh_aas + n_tmh_aas
  testthat::expect_equal(
    n_aas,
    sum(nchar(protr::readFASTA(fasta_filename)))
  )
  perc_tmh_aas <- n_tmh_aas / (n_tmh_aas + n_non_tmh_aas)
  n_tmhs <- sum(tally_tmhs(topology)$n_tmhs)

  tibble::tibble(
    n_aas = n_aas,
    n_proteins = n_proteins,
    n_non_tmh_aas = n_non_tmh_aas,
    n_tmh_aas = n_tmh_aas,
    perc_tmh_aas = perc_tmh_aas,
    n_tmhs = n_tmhs
  )
}

```

```{r}
t1 <- show_properties(fasta_filename = "proteome.fasta")
knitr::kable(t1)
```

```{r}
t2 <- show_properties(fasta_filename = "UP000464024.fasta")
knitr::kable(t2)
```