### 2020-06-08 p_bind_per_hydrophobicity/p_bind_per_hydrophobicity.Rmd
Try to render with Pandoc

---
title: "p_bind"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Calculate the relation between hydrophobicity
# and chance to bind to MHC-I.
#
# H0: there is no relation
# H1: higher hydrophobicity equals higher chance to bind
#
suppressMessages(library(ggplot2))
suppressMessages(library(dplyr))
suppressMessages(library(testthat))

plot_is_tmh_vs_binds_mhc1 <- function(
  df,
  png_filename = tempfile(fileext = ".png")
) {
  df_tally <- df %>%
    dplyr::group_by(is_tmh, binds_mhc1) %>%
    dplyr::tally()
  ggplot2::ggplot(df_tally, ggplot2::aes(is_tmh, binds_mhc1, fill = n)) +
    ggplot2::geom_tile() +
    ggplot2::geom_text(ggplot2::aes(label = n)) +
      ggplot2::scale_fill_gradient(low = "white", high = "red") +
      ggplot2::scale_x_discrete() +
      ggplot2::scale_y_discrete() +
      ggplot2::xlab("Is TMH? Left = no, right = yes") +
      ggplot2::ylab("Binds to MHC-I? Down = no, up = yes") +
      ggplot2::ggsave(png_filename, width = 7, height = 7)
}

plot_is_tmh_vs_binds_mhc2 <- function(
  df,
  png_filename = tempfile(fileext = ".png")
) {
  df_tally <- df %>%
    dplyr::group_by(is_tmh, binds_mhc2) %>%
    dplyr::tally()
  ggplot2::ggplot(df_tally, ggplot2::aes(is_tmh, binds_mhc2, fill = n)) +
    ggplot2::geom_tile() +
    ggplot2::geom_text(ggplot2::aes(label = n)) +
      ggplot2::scale_fill_gradient(low = "white", high = "red") +
      ggplot2::scale_x_discrete() +
      ggplot2::scale_y_discrete() +
      ggplot2::xlab("Is TMH? Left = no, right = yes") +
      ggplot2::ylab("Binds to MHC-II? Down = no, up = yes") +
      ggplot2::ggsave(png_filename, width = 7, height = 7)
}

plot_binds_mhc1_vs_binds_mhc2 <- function(
  df,
  png_filename = tempfile(fileext = ".png")
) {
  df_tally <- df %>%
    dplyr::group_by(binds_mhc1, binds_mhc2) %>%
    dplyr::tally()
  ggplot2::ggplot(df_tally, ggplot2::aes(binds_mhc1, binds_mhc2, fill = n)) +
    ggplot2::geom_tile() +
    ggplot2::geom_text(ggplot2::aes(label = n)) +
      ggplot2::scale_fill_gradient(low = "white", high = "red") +
      ggplot2::scale_x_discrete() +
      ggplot2::scale_y_discrete() +
      ggplot2::xlab("Binds to MHC-I?? Left = no, right = yes") +
      ggplot2::ylab("Binds to MHC-II? Down = no, up = yes") +
      ggplot2::ggsave(png_filename, width = 7, height = 7)
}

plot_hydrophobicity_vs_is_tmh <- function(
  df,
  png_filename = tempfile(fileext = ".png")
) {
  df <- tibble::as_tibble(df)

  df_tally <- df %>%
    dplyr::mutate(bin = trunc(hydrophobicity * 2)) %>%
    dplyr::group_by(bin, is_tmh) %>%
    dplyr::tally()
  ggplot2::ggplot(df_tally, ggplot2::aes(bin, is_tmh, fill = n)) +
    ggplot2::geom_tile() +
    ggplot2::geom_text(ggplot2::aes(label = n)) +
      ggplot2::scale_fill_gradient(low = "white", high = "red") +
      ggplot2::scale_y_discrete() +
      ggplot2::xlab("Hydrophobicity x2") +
      ggplot2::ylab("Is TMH? Down = no, up = yes") +
      ggplot2::ggsave(png_filename, width = 7, height = 7)
}

plot_hydrophobicity_vs_binds_mhc1 <- function(
  df,
  png_filename = tempfile(fileext = ".png")
) {
  df <- tibble::as_tibble(df)
  df_tally <- df %>%
    dplyr::mutate(bin = trunc(hydrophobicity * 2)) %>%
    dplyr::group_by(bin, binds_mhc1) %>%
    dplyr::tally()
  ggplot2::ggplot(df_tally, ggplot2::aes(bin, binds_mhc1, fill = n)) +
    ggplot2::geom_tile() +
    ggplot2::geom_text(ggplot2::aes(label = n)) +
      ggplot2::scale_fill_gradient(low = "white", high = "red") +
      ggplot2::scale_y_discrete() +
      ggplot2::xlab("Hydrophobicity x2") +
      ggplot2::ylab("Binds to MHC-I? Down = no, up = yes") +
      ggplot2::ggsave(png_filename, width = 7, height = 7)
}

plot_hydrophobicity_vs_binds_mhc2 <- function(
  df,
  png_filename = tempfile(fileext = ".png")
) {
  df <- tibble::as_tibble(df)

  df_tally <- df %>%
    dplyr::mutate(bin = trunc(hydrophobicity * 2)) %>%
    dplyr::group_by(bin, binds_mhc2) %>%
    dplyr::tally()
  ggplot2::ggplot(df_tally, ggplot2::aes(bin, binds_mhc2, fill = n)) +
    ggplot2::geom_tile() +
    ggplot2::geom_text(ggplot2::aes(label = n)) +
      ggplot2::scale_fill_gradient(low = "white", high = "red") +
      ggplot2::scale_y_discrete() +
      ggplot2::xlab("Hydrophobicity x2") +
      ggplot2::ylab("Binds to MHC-II? Down = no, up = yes") +
      ggplot2::ggsave(png_filename, width = 7, height = 7)
}

raw_data <- "p_bind_per_hydrophobicity.csv"
expect_true(file.exists(raw_data))
df <- read.csv(raw_data)

plot_is_tmh_vs_binds_mhc1(df = df, png_filename = "is_tmh_vs_binds_mhc1.png")
plot_is_tmh_vs_binds_mhc2(df = df, png_filename = "is_tmh_vs_binds_mhc2.png")
plot_binds_mhc1_vs_binds_mhc2(df = df, png_filename = "binds_mhc1_vs_binds_mhc2.png")
plot_hydrophobicity_vs_is_tmh(df = df, png_filename = "hydrophobicity_vs_is_tmh.png")
plot_hydrophobicity_vs_binds_mhc1(df = df, png_filename = "hydrophobicity_vs_binds_mhc1.png")
plot_hydrophobicity_vs_binds_mhc2(df = df, png_filename = "hydrophobicity_vs_binds_mhc2.png")
```