### 2021-03-11 ncbi_peregrine_results/ncbi_counts.tex
Add constants

\begin{table}

\caption{\label{tab:ncbi_counts}Amounts. map = MAP = membrane associated protein. tmp = TMP = transmembrane protein. in\_tmh = in transmembrane helix (TMH) of TMP. in\_sol = in soluble region of TMP. perc\_tmh = percentage transmembrane helix (TMH). }
\centering
\begin{tabular}[t]{l|r|r|r|r|r|r}
\hline
what & raw & all\_proteins & map & tmp & in\_tmh & in\_sol\\
\hline
Number of variations & 60931 & 37831 & 16623 & 21208 & 3803 & 17405\\
\hline
Number of unique variations & 60544 & 37630 & 16606 & 21024 & 3789 & 17235\\
\hline
Number of unique SNPs & NA & 9621 & 4219 & 6026 & 1140 & 4936\\
\hline
Number of unique gene names & 953 & 911 & 457 & 605 & 325 & 590\\
\hline
Number of unique protein names & 5163 & 4780 & 2227 & 2553 & 1280 & 2467\\
\hline
Percentage TMH & NA & 10 & 0 & 19 & 26 & 18\\
\hline
\end{tabular}
\end{table}