### 2021-03-18 ncbi_peregrine_results/snp_stats_per_spanner_multi.tex

\begin{table}

\caption{\label{tab:snp_stats_per_spanner_multi}Statistics for the multi-spanners. p = p value. n = number of SNPs in multi-spanners. n\_success = number of SNPs found in TMHs of multi-spanners (dashed blue line). E(n\_success) = expected number of SNPs to be found in TMHs of multi-spanners  (dashed red line). }
\centering
\begin{tabular}[t]{l|l}
\hline
parameter & value\\
\hline
p & 8.315841e-12\\
\hline
n & 13022\\
\hline
n\_success & 3351\\
\hline
E(n\_success) & 3678.406\\
\hline
\end{tabular}
\end{table}