### 2021-03-18 ncbi_peregrine_results/ncbi_counts_2.tex

\begin{table}

\caption{\label{tab:ncbi_counts}Amounts. single\_in\_tmh = in transmembrane helix of single-spanner. single\_in\_sol = in soluble region of single-spanner. multi\_in\_tmh = in transmembrane helix of multi-spanner. multi\_in\_sol = in soluble region of multi-spanner. }
\centering
\begin{tabular}[t]{l|r|r|r|r}
\hline
what & single\_in\_tmh & single\_in\_sol & multi\_in\_tmh & multi\_in\_sol\\
\hline
Number of variations & 452 & 7734 & 3351 & 9671\\
\hline
Number of unique variations & 451 & 7733 & 3338 & 9502\\
\hline
Number of unique SNPs & 160 & 2393 & 994 & 2762\\
\hline
Number of unique gene names & 96 & 282 & 243 & 344\\
\hline
Number of unique protein names & 304 & 1032 & 976 & 1435\\
\hline
Percentage TMH & 11 & 5 & 35 & 26\\
\hline
\end{tabular}
\end{table}