### 2021-04-12 bbbq_article.tex
Updates from Overleaf

\begin{abstract}

Cytolytic T cell responses are predicted to be biased towards membrane proteins. The peptide-binding grooves of most haplotypes 
of histocompatibility complex class I (MHC-I) are relatively hydrophobic,  therefor peptide fragments derived from human transmembrane helices (TMHs) are predicted to be presented more often as would be expected 
based on their abundance in the proteome. However, the physiological reason of why membrane proteins might be over-presented is unclear. 
In this study, we show that the over-presentation of TMH-derived peptides is general, as it is predicted for bacteria and viruses 
and for both MHC-I and MHC-II. Moreover, we show that TMHs are evolutionarily more conserved, because single nucleotide polymorphisms (SNPs) are present relatively less frequently in TMH-coding chromosomal regions compared to regions coding for extracellular and cytoplasmic protein regions. Thus, our findings suggest that both cytolytic and helper T cells respond more to membrane proteins, because these are evolutionary more conserved. We speculate that TMHs therefor are less prone to escape mutations that enable pathogens to evade T cell responses.

\end{abstract}

{\bf Keywords:} antigen presentation, membrane proteins, bioinformatics, 
adaptive immunity, transmembrane domain, transmembrane helix, 
epitopes, T lymphocyte, MHC-I, MHC-II, evolutionary conservation

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% \paragraph{Immune response}

Our immune system fights diseases and infections from pathogens, 
such as fungi, bacteria or viruses. 
An important part of the acquired immune response, 
that develops specialized and more specific recognition of pathogens 
than the innate immune response, 
are T cells which recognize peptides, called epitopes, derived from antigenic proteins presented on Major Histocompatibility Complexes (MHC) class I and II on the cell surface.

% \paragraph{Classification of HLA}

The MHC proteins are heterodimeric complexes encoded by the
HLA (Human Leukocyte Antigens) genes.
In humans, the peptide binding groove of MHC-I is made by only the alpha subunit. There are three classical forms of MHC-I, hallmarked by a highly polymorphic alpha chain called HLA-A, HLA-B and HLA-C, that all present epitopes to cytolytic T cells. 
For MHC-II, both the alpha and the beta chains contribute to the peptide binding groove. There are three classical forms of MHC-II as well, called HLA-DR, HLA-DQ and HLA-DP, that all present epitopes to helper T cells.
Each MHC complex can present a subset of all possible peptides.
For example, HLA-A and HLA-B have no overlap in which
epitopes they bind (\cite{lund2004definition}).
Moreover, the HLA genes of humans are highly polymorphic, with hundreds 
to thousands of different alleles, 
and each different HLA allele is called 
an MHC haplotype and presents a different subset of peptides (\cite{marsh2010nomenclature}).

% \paragraph{HLAs increase detection range}

Humans mostly express two haplotypes per MHC form, one from the parental and one from the maternal chromosome, and therefore an individual's immune system detects 
only a fraction of all possible peptide fragments. 
However, at the population level, the coverage of pathogenic peptides that are detected 
is very high, because of the highly polymorphic MHC genes.
It is therefor believed that MHC polymorphism improves immunity at the population level, 
as mutations in a protein that disrupt a particular MHC presentation at the individual level, 
so-called escape mutations, 
will not affect MHC presentation for all haplotypes present in the population  (\cite{sommer2005importance}).

% \paragraph{Epitope prediction}

Many studies are aimed at identifying the repertoire of epitopes that are presented in any MHC haplotype and determining which epitopes will result in an immune response, 
as this will for instance aid the design of vaccines. 
These studies have led to the development of prediction algorithms 
that allow for very reliable \emph{in silico} predictions 
of the binding affinities of peptides
(\cite{larsen2010identification,schellens2008unanticipated,tang2011genome}).
For example, \cite{tang2011genome} found that, 
of the 432 peptides that were predicted to bind to MHC,
86\% were experimentally confirmed to do so.

% \paragraph{TMHs}

Using these prediction algorithms, 
we recently predicted that peptides derived 
from transmembrane helices (TMHs) 
will be more frequently presented by MHC-I 
than expected based on their abundance (\cite{bianchi2017}). 
Moreover, we showed that some well-known immunodominant peptides stem from TMHs. 
This over-presentation is attributed to the fact 
that the peptide-binding groove of most MHC-I haplotypes 
is relatively hydrophobic, 
and therefore hydrophobic peptides have a higher affinity 
to bind than more hydrophilic peptides.

TMHs are hydrophobic 
as they need to span the hydrophobic lipid bilayer of cellular membranes.
They consist of an alpha helix of, on average, 23 amino acids in length. 
TMHs can also be predicted with high accuracy from a protein sequence 
by bioinformatics approaches (\cite{krogh2001predicting,bianchi2017,kall2004combined,arai2004conpred,jones2007improving,klammer2009metatm,wang2019efficient}),
for example, \cite{jones2007improving} found that,
from 184 transmembrane proteins (TMPs) with known topology, 
80\% of the TMH predictions replicated this finding.

TMHs are common structures in the proteins of humans and microbes. 
Different TMH prediction tools estimate that 15-39\% of all proteins 
in the human proteome contain at least one TMH (\cite{ahram2006estimation}).
However, the physiological reason why peptides derived from TMHs 
would be presented more often than peptides 
stemming from soluble (i.e., extracellular or cytoplasmic) protein regions is unknown. 
We hypothesized that the presentation of 
TMH residues is evolutionary selected for, 
because TMHs are less prone to undergo escape mutations. 
One reason to expect such a reduced 
variability (and hence evolutionary conservation) in TMHs, 
is that these are restricted in their evolution 
by the functional requirement to span a lipid bilayer. 
Due to this requirement, 
many of the amino acids genuinely present in TMHs 
are limited to the ones with hydrophobic side chains (\cite{jones1994model}).
Therefore, we speculated that the TMHs of pathogens 
might have a lower chance to develop escape mutations, 
as many mutations will result in a dysfunctional TMH 
and render the protein inactive.

This study had two objectives. 
First, we aimed to generalize our findings by predicting the presentation of peptides 
from different kingdoms of life and for both MHC-I and -II. 
From these \emph{in silico} predictions, we conclude that TMH-derived
epitopes are presented more often than expected by chance,
in a human, viral and bacterial proteome, 
and for most haplotypes of both MHC-I and II. 
We confirmed the presentation of TMH-derived peptides 
by re-analysis of peptide elution studies. 
Second, we tested our hypothesis that TMHs 
are more evolutionary conserved than soluble protein regions. 
Our analysis of human single nucleotide polymorphisms (SNPs) showed 
that random point mutations are indeed less likely
to occur within TMHs. 
These findings strengthen the emerging notion 
that TMHs are important for the T cell branch of the adaptive immune system, 
and hence are of  
overlooked importance in vaccine development.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Methods}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Predicting TMH epitopes}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

To predict how frequently epitopes overlapping with TMHs are presented,
a similar analysis strategy was applied as described in \cite{bianchi2017} 
for several haplotypes of both MHC-I and MHC-II, 
and for a human, viral and bacterial proteome.
To summarize, for each proteome, 
all possible 9-mers (for MHC-I) or 14-mers (MHC-II) were derived. 
For each of these peptides, we determined if it overlapped with a predicted 
TMH and if it was predicted to bind to each haplotype.

For MHC-I, 9-mers were used, as this is the length most frequently presented in MHC-I and was used in our earlier study (\cite{bianchi2017}). 
For MHC-II, 14-mers were used, 
as these are the most frequently occurring epitope length (\cite{bergseng2015different}).
A human (UniProt ID UP000005640\_9606), 
viral (SARS-CoV-2, UniProt ID UP000464024) 
and bacterial (\emph{Mycobacterium tuberculosis}, UniProt ID UP000001584) 
reference proteome was used. 
\verb;TMHMM; (\cite{krogh2001predicting}) was used to predict the topology 
of the proteins within these proteomes.
To predict the affinity of an epitope to a certain MHC haplotype,
 \verb;EpitopePrediction; (\cite{bianchi2017}) for MHC-I 
and \verb;MHCnuggets; (\cite{shao2020high}) for MHC-II was used.
The 13 MHC-I haplotypes used in this study are the same as 
used in the previous study (\cite{bianchi2017}).
For MHC-II, haplotypes were selected with a phenotypic frequency 
of at least 14\% in the human population (\cite{greenbaum2011functional}),
resulting in 21 MHC-II haplotypes.

\geert{wat text gewist want dit is de methods sectie}
In previous work, it was found that the over-presentation of TMH-derived peptides can be
explained from the hydrophobicity of the MHC-I binding cleft (\cite{bianchi2017}).
Here, a similar analysis was applied, by correlating the percentage of predicted TMH-derived epitopes versus the mean hydrophobicity of all 
peptides.

This study differs in one important aspect from our previous work (\cite{bianchi2017}). 
The definition of a binder differs from \cite{bianchi2017}:
in the current study, a peptide is called a binder if, for a certain haplotype, 
any of its 9-mer or 14-mer peptides have an IC50 value in the lowest 2\% of 
all peptides within a \emph{proteome} 
(see tables \ref{tab:ic50_binders_mhc1} and \ref{tab:ic50_binders_mhc2} for values), 
whereas the previous study defined
a binder as having an IC50 in the lowest 2\% 
of the peptides within a \emph{protein}.
% See https://github.com/richelbilderbeek/bianchi_et_al_2017/blob/72e6755a31d400158368509fd80a41e984677ab1/predict-binders.R#L17
This revised definition precludes bias of proteins 
that give rise to no or only very few MHC epitopes.
To verify that the results are similar, a side by side comparison was performed shown in the supplementary materials. \geert{refereer hier naar supplementary figuur}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Peptide elution studies}\label{subsec:elution_studies}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

To obtain experimental evidence that epitopes derived from TMHs are presented in MHC, peptide elution studies for MHC-I (\cite{schellens2015comprehensive}) 
and MHC-II (\cite{bergseng2015different}) were reanalyzed.
For each of the detected epitopes,
its possible location(s) in 
a human reference proteome,
with UniProt ID UP000005640\_9606, was mapped.
For the epitopes that were present in the proteome exactly once,
the topology of the proteins in which these epitopes were located was predicted using both \verb;TMHMM; (\cite{krogh2001predicting}) and \verb;PureseqTM; (\cite{wang2019efficient}). 
From this topology, we determined if the epitope
overlapped with a TMH.

The full analysis can be found
at \url{https://github.com/richelbilderbeek/bbbq_article_issue_157}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Evolutionary conservation of TMHs}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% \paragraph{Introduction}

To determine the evolutionary conservation of TMHs,
human single nucleotide polymorphisms (SNPs) were first collected
that resulted in a single amino acid substitution,
and we then determined if this substitution occurred within a predicted TMH or not.

% \paragraph{Data}

As a data source, multiple
NCBI (\url{https://www.ncbi.nlm.nih.gov/}) databases were used: the \emph{dbSNP} (\cite{sherry2001dbsnp}) database,
which contains 650 million 
catalogued non-redundant humane variations (called RefSNPs,
\url{https://www.ncbi.nlm.nih.gov/snp/docs/RefSNP_about/}), and the databases \emph{gene} (for gene names, \cite{brown2015gene})
and \emph{protein} (for proteins sequences, \cite{sayers2010database}).

% \paragraph{Pipeline}

The first query was a call to the \emph{gene} database for the 
term 'membrane protein' (in all fields) 
for the organism \emph{Homo sapiens}.
This resulted in 1,077 gene IDs (on December 2020).
% 1077 gene IDs is correct for December 2020.
% At 2021-03-01, one will get 1130 gene IDs.
% Also, one of the gene IDs that was valid back then,
% has been obsoleted.
The next query was a call to the \emph{gene} database 
to obtain the gene names from the gene IDs.
Per gene name, the \emph{dbSNP} NCBI database was queried for 
variations associated with the gene name. 
As the NCBI API constrains its users to three calls per second
(to assure fair use), we had to limit the extent of our analysis.

The number of SNPs was limited to the first 250 variations per gene,
resulting in $\approx$61k variations.
Only variations that result in a SNP for
a single amino acid substitution were analyzed, resulting in $\approx$38k SNPs.
The exact amounts can be found in the supplementary materials,
tables \ref{tab:ncbi_counts_1} and \ref{tab:ncbi_counts_2}.

% \paragraph{Selection of SNPs}
%
SNPs were picked based on ID number, which is linked to their discovery date. To verify that these ID numbers are unrelated to SNP positions, the relative positions of all analyzed SNPs in a protein were determined. This analysis showed no positional bias of the SNPs, as shown in \geert{dit is supplement toch?} supplementary figure \ref{fig:snp_rel_pos}.

Per SNP, the \emph{protein} NCBI database was queried for the
protein sequence. For each protein sequence, the protein topology was determined using \verb;PureseqTM;. Using these predicted protein topologies, the SNPs were scored to be located within or outside TMHs.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{TMH-derived peptides are predicted to be over-presented in MHC-I}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Figure \ref{fig:bbbq_1_smart_results} shows the predicted presentation of TMH-derived peptides in MHC-I,
for a human, viral and bacterial proteome.
Per MHC-I haplotype, it shows the percentage of binders that overlap with a TMH 
with at least one residue.
The horizontal line shows the expected percentage of TMH-derived epitopes 
that would be presented, if TMH-derived epitopes would be presented just as 
likely as epitopes derived from soluble regions.
For 11 out of 13 MHC-I haplotypes, TMH-derived epitopes are predicted to be presented more often 
than the null expectation, for a human and bacterial proteome.
For the viral proteome, 12 out of 13 haplotypes present
TMH-derived epitopes more often than expected by chance.
The extent of the over-presentation between the different haplotypes 
is similar for the probed proteomes, 
which strengthens our previous conclusion (\cite{bianchi2017}) 
that the hydrophobicity of the MHC-binding groove 
is the main factor responsible for the predicted over-presentation 
of TMH-derived peptides.  \frans{zoals eerder HLA-C mist en ook HLA-DP voor de nu gekozen haplotypes mis ik de rationale waarom we deze hebben genomen? volgens mij had ik een artikel gestuurd met de meest voorkomende HLA-DP haplotypes. De rationale voor HLA-C is dat de expressie hier naar verhousding heel laag van is en dat dus een minder explecitie rol speelt in antigen presentatie} \geert{in methods section geeft Richel al rationale voor de keuze van de haplotypes, waarin hij schrijft dat dit gebasseerd is op onze eerdere studie (voor MHC-I) en de frequentie van de haplotypes (voor MHC-II). Dit lijkt mij voldoende.}

\begin{figure}
  \centering
  \begin{subfigure}[t]{0.9\textwidth}
    \centering
    \caption{}
    \includegraphics[width=\linewidth]{bbbq_1_smart_results/fig_f_tmh_2_panel.png}
    \label{fig:bbbq_1_smart_results}
  \end{subfigure}

  \vfill

  \begin{subfigure}[t]{0.4\textwidth}
    \centering
    \caption{}
    \includegraphics[width=\linewidth]{bbbq_1_smart_results/fig_hydrophobicity_mhc1.png}
    \label{fig:hydrophobicity_1}
  \end{subfigure}  
  \hfill
  \begin{subfigure}[t]{0.4\textwidth}
    \centering
    \caption{}
    \includegraphics[width=\linewidth]{bbbq_1_smart_results/fig_hydrophobicity_mhc2.png}
    \label{fig:hydrophobicity_2}
  \end{subfigure}  
  \caption{ \textbf{Over-presentation of TMH-derived epitopes on most MHC-I and -II haplotypes} \geert{een figuur heeft altijd een titel met de conclusie. De legenda zelf is feitelijke opsomming en biedt geen (of weinig) interpretatie}
    \textbf{(A)} 
    The percentage of epitopes for MHC-I and -II haplotypes that are predicted to 
    overlap with TMHs for the proteomes of SARS-CoV-2 (top row), human (middle 
    row) and \emph{M. tuberculosis} (bottom row).
    The red lines indicate the percentages as expected by chance.
    See supplementary tables \ref{tab:tmh_binders_mhc1} and \ref{tab:tmh_binders_mhc2}
    for the exact TMH and  epitope counts.
    \textbf{(B-C)}
    Correlation between the percentages of predicted TMH-derived epitopes
    and the hydrophobicity score of all predicted epitopes for MHC-I \textbf{(B)}
    and MHC-II haplotypes \textbf{(C)}. Red curve: linear regression analysis. Labels are shorthand for the HLA haplotypes,
    see Table \ref{tab:haplotype_abbreviations} for the names. \frans{wat is de reden dat nr 15 en 8 er buiten vallen? de rest is vrijwel linear)
  }}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{TMH-derived peptides are predicted to be over-presented in MHC-II}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

We next wondered if the over-representation of TMH-derived peptides 
would also be present for MHC-II. 
Figure \ref{fig:bbbq_1_smart_results} shows the percentages of MHC-II epitopes 
predicted to be overlapping with TMHs for our human, viral and bacterial proteomes.
We found that TMH-derived peptides are over-presented in all
of the 21 MHC-II haplotypes, 
for a human, bacterial and viral proteome,
except for \verb;HLA-DRB3*0101; in \emph{M. tuberculosis}.
See supplementary table \ref{tab:tmh_binders_mhc2} 
for the exact TMH and epitope counts.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{The over-presentation of TMH-derived peptides is caused by the hydrophobicity of the MHC peptide binding groove}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

For MHC-I, we previously showed that the over-presentation of TMH-derived peptides is caused by the hydrophobicity of the peptide binding grooves \cite{bianchi2017}. Figures \label{fig:hydrophobicity_1} and \label{fig:hydrophobicity_2} \frans{de verwijzing werkt denk ik niet}
show the extent of over-presentation
of TMH-derived epitopes as a function of the hydrophobicity preference score for the different haplotypes.
An assumed linear correlation explains 88\% of the variability in MHC-I.
For MHC-II, 62\% of the variability is explained by hydrophobicity. \frans{dit is dus omdat 2 haplotypes buiten de boot vallen} This strengthens our previous finding \cite{bianchi2017} and indicates that TMH-derived peptides are over-presented because the peptide binding grooves of most MHC-I and -II haplotypes are relatively hydrophobic.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Experimental validation of presentation of TMH-derived peptides}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}[!htbp]
  \centering
  \includegraphics[width=0.5\textwidth]{bbbq_article_issue_157/results.png}
  \caption{ \textbf{MHC-I and -II presentation of TMH-derived peptides in peptide elution studies.}
    TMH-derived epitopes are presented \emph{in vivo},
    for both MHC-I and MHC-II, for two topology prediction
    programs. \geert{legenda is veel te summier. Wat is precies geplot? ook studies van de data citeren.}
  }
  \label{fig:elution}
\end{figure}

To obtain experimental confirmation that peptides stemming from TMHs 
are presented in MHC-I and MHC-II,
two peptide elution studies (\cite{schellens2015comprehensive} and \cite{bergseng2015different} for MHC-I and MHC-II respectively) 
were reanalyzed.
Figure \ref{fig:elution} shows the percentages of epitopes derived 
from TMHs
found in the MHC-I and MHC-II elution 
studies, \geert{is dit van totale peptides, of van humane peptides? Uiteraard van humane peptides gebruiken, want de rest heb je niet gemapped. Dat moet dan wel hier gezegd worden. Daarnaast: zijn de MHC-I en -II haplotypes bekend in deze studies?}
for the two topology prediction tools \verb;TMHMM; (\cite{krogh2001predicting}) and \verb;PureseqTM; (\cite{wang2019efficient}). 
Regardless of the prediction tool, 
at least 100 epitopes were predicted to be derived from a TMH for each condition. 
From these findings, it is robustly predicted that
epitopes derived from TMHs are presented in both MHC-I and MHC-II.
See the supplementary materials (table \ref{tab:elution}) for the exact values.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Human TMHs are evolutionarily conserved}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}
  \centering
  \begin{subfigure}[t]{0.45\textwidth}
    \centering
    \caption{}
    \includegraphics[width=\linewidth]{ncbi_peregrine_results/fig_conservation.png}
    \label{fig:conservation}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.45\textwidth}
    \centering
    \caption{}
    \includegraphics[width=\linewidth]{ncbi_peregrine_results/fig_f_snps_found_and_expected.png}
    \label{fig:f_snps_found_and_expected}
  \end{subfigure}

   \vfill

  \begin{subfigure}[t]{0.45\textwidth}
    \centering
    \caption{}
    \includegraphics[width=\linewidth]{ncbi_peregrine_results/fig_f_tmh_ncbi.png}
    \label{fig:f_tmh_ncbi}
  \end{subfigure}

  \caption{ \textbf{Evolutionary conservation of human TMHs.}
    \textbf{(A)} 
    The number of SNPs in TMHs as expected by chance (left bar) 
    and found in the dbSNP database (right bar).
    Percentages show the relative conservation
    of SNPs in TMHs found.
    \textbf{(B)}
    Percentage of SNPs found in TMHs.
    Each point shows for one protein the predicted percentage of
    TMH ($x$-axis) and the observed occurrence of SNPs being located
    within a TMH ($y$-axis).
    The dashed diagonal line shows the line of equality (i.e.,
    equal conservation of TMHs and soluble protein regions). The red line shows the linear regression analysis and the grey shaded area the 95\% confidence interval. \geert{please correct last sentence}
    \textbf{(C)}
    Distribution of the percentages of TMH in the TMPs used in this study.
  }
\end{figure}

\begin{figure}
  \centering
  \begin{subfigure}[t]{0.45\textwidth}
    \centering
    \caption{}
    \includegraphics[width=\linewidth]{ncbi_peregrine_results/fig_conservation_per_spanner.png}
    \label{fig:conservation_per_spanner}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.45\textwidth}
    \centering
    \caption{}
    \includegraphics[width=\linewidth]{ncbi_peregrine_results/fig_f_tmh_ncbi_per_spanner.png}
    \label{fig:f_tmh_ncbi_per_spanner}
  \end{subfigure}

  \vfill

  \begin{subfigure}[t]{\textwidth}
    \centering
    \caption{}
    \includegraphics[width=\linewidth]{ncbi_peregrine_results/fig_f_snps_found_and_expected_per_spanner.png}
    \label{fig:f_snps_found_and_expected_per_spanner}
  \end{subfigure}
  \caption{ \textbf{Membrane proteins with multiple TMHs are evolutionary more conserved than proteins with only a single TMH.}
      \textbf{(A)} 
      The number of SNPs in TMHs as expected by chance 
      and observed in the dbSNP database, 
      for TMPs with one TMH (single-spanners) and multiple TMHs (multi-spanners).
      Percentages show the relative conservation
      of SNPs in TMHs found.
      \textbf{(B)} 
      Distribution of the proportion of amino acids residing
      in the plasma membrane. 
      \textbf{(C)} 
      Percentage of SNPs found in TMPs predicted to have only a single
      (left) or multiple (right) TMHs.
        Each point shows for one protein the predicted percentage of TMH ($x$-axis) and the observed occurrence of SNPs being located within a TMH ($y$-axis). The dashed diagonal lines show the line of equality (i.e., equal conservation of TMHs and soluble protein regions). The red lines shows the linear regression analyses and the grey shaded areas the 95\% confidence intervals. \geert{please correct last sentence}
     }
\end{figure}

We addressed the question whether there is an evolutionary advantage in presenting TMHs.
We determined the conservation of TMHs 
by comparing the occurrences of SNPs located in TMHs or soluble protein regions 
for the genes coding for membrane proteins.
We obtained 911 unique gene names associated with the phrase 'membrane protein',
which are genes coding for both membrane-associated proteins (MAPs, which have no TMH) and 
transmembrane proteins (TMPs, which have at least one TMH).
These genes are linked to 4,780 protein isoforms, 
of which 2,553 are predicted to be TMPs and 
2,237 proteins are predicted to be MAPs.
We obtained 37,630 unique variations, 
of which 9,621 are SNPs that resulted in a straightforward amino acids substitution, 
of which 6,062 were located in predicted TMPs.
See tables \ref{tab:ncbi_counts_1} and \ref{tab:ncbi_counts_2} 
for the detailed numbers and distributions of SNPs.

Per protein, we calculated two percentages: 
(1) the percentage of the total protein predicted to be TMHs, 
and (2) the percentage of SNPs located within these predicted TMHs.
Each percentage pair was plotted in figure \ref{fig:f_snps_found_and_expected}.
The proportion of SNPs found in TMHs varied from 
none (i.e. all SNPs were in
soluble regions) to all.
To determine if SNPs were randomly distributed over the protein, we performed a linear regression analysis,
and added a 95\% confidence interval on this regression.
This linear fit nearly goes through the origin and has a slope
below the line of equality,
which shows that less SNPs are found in TMHs than expected by chance.

We determined the probability to find the observed amount
of SNPs in TMHs by chance, i.e., when assuming SNPs occur 
just as likely in soluble domains as in TMHs.
We used a binomial Poisson distribution, 
where the number of trails ($n$) equals the number of SNPs, 
which is 21,208. 
The probability of success for the $i$th TMP ($p\_i$), 
is the percentage of residues within a TMH per TMP. 
These percentages are shown as a histogram 
in figure \ref{fig:f_tmh_ncbi}. 
The expected number of SNPs expected to be found in 
TMHs by chance equals $\sum{p} \approx 4,141$.
As we observed 3,803 SNPs in TMHs, 
we calculated the probability of having that amount or less successes.
We used the type I error cut-off value of $\alpha = 2.5\%$.
The chance to find, within TMHs, this amount or less SNPs 
equals $6.8208 \cdot 10^{-11}$.
We determined the relevance of this finding, by
calculating how much less SNPs are found in TMHs,
when compared to soluble regions, which is the
ratio between the number of SNPs found in TMHs
versus the number of SNPs as expected by chance.
In effect, per 1000 SNPs found in soluble protein domains, 
one finds 918 SNPs in TMHs,
as depicted in figure \ref{fig:conservation}.

We split this analysis for TMPs containing only a single TMH (so-called single-membrane spanners) and TMPs containing multiple TMHs (multi-membrane spanners). 
We hypothesized that single-membrane spanners are less conserved than multi-membrane spanners,
because multi-membrane spanners
might have protein-protein interactions between their TMHs, 
for example to accommodate active sites, and 
thus might have additional structural constraints.
From the split data, we did the same analysis as for the total TMPs.
Figure \ref{fig:f_snps_found_and_expected_per_spanner} 
shows the percentages of TMHs for individual proteins as a function of the
percentage of SNPs located in TMHs.
For both single- and multi-spanners, a linear regression shows that less
SNPs are found in TMHs, than expected by chance.

We also determined the probability to find the 
observed amount of SNPs  by chance in single- and multi-spanners.
For single-spanners, we found 452 SNPs in TMH, where
$\approx462$ were expected by chance. 
The chance to observe this or a lower number by chance is 
$0.319$. As this chance was higher than our $\alpha = 0.025$,
we consider this no significant effect.
For the single-spanners, we found 3,351 SNPs in TMH, where \frans{er staat 2keer single spanners dit zijn. toch de multispanners?}
$\approx3,678$ were expected by chance. 
The chance to observe this or a lower number by chance is 
$8.315841 \cdot 10^{-12}$, 
which means this number is significantly less as explained by variation.

Also, for single- and multi-spanners, 
we determined the relevance of this finding by
calculating how much less SNPs are found in TMHs
when compared to soluble regions,
as depicted in figure \ref{fig:conservation_per_spanner}.
In effect, per 1,000 SNPs found in soluble protein domains, 
one finds 978 SNPs in TMHs 
of single-spanners (although this is not significant)
and 911 SNPs in TMHs of multi-spanners.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% \paragraph{General}

Epitope prediction is important to understand the immune system 
and for the design of vaccines.
In this study, we provide evidence that epitopes
derived from TMHs are a major but overlooked source of MHC epitopes. 
Our bioinformatics predictions indicate that TMH-derived epitopes 
are presented to both cytolytic and helper T cells more often than expected by chance, 
regardless of the organism. 
Moreover, reanalysis of peptide elution studies confirmed the presentation of TMH-derived epitopes.
Finally, our SNP analysis shows that TMHs are evolutionary more conserved 
than solvent-exposed protein regions.

\subsection{Mechanism of MHC presentation of TMH-derived epitopes}
% \paragraph{TMH presentation in vivo}

Although our data show that 
TMH-derived epitopes are presented in MHC-I and MHC-II, 
the molecular mechanisms of how integral membrane proteins are processed 
for MHC presentation are largely unknown (\cite{bianchi2017}). 
Most prominently, the fundamental principles of 
how TMHs are extracted from the hydrophobic lipid environment 
into the aqueous vacuolar lumen, 
and their prior or subsequent proteolytic processing are unresolved, as we discussed previously for MHC-I presentation \cite{bianchi2017}. 
\geert{
  Hier toevoegen wat probleem is MHC-II. Maxim had goed idee voor intraluminal vesicles in MHC-II loading.
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Experimental evidence for MHC presentation of TMH-derived epitopes}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\geert{Deze gehele subsection kan wat mij betreft wel weg (of naar de 'minor discussion' in de supplement). Het is deels een herhaling van resultaten (1e en 3e paragraaf). DRiPs in paragraaf 2 kan echt niet alles verklaren, want MHC-II is opname van heel antigen. In parafraag 4, zal de methodological bias zal alleen maar resulteren in het onderschatten van presentatie van TMHs, en is dus ook niet echt relevant hier.}
% \paragraph{Over-presentation in vivo in MHC-I and MHC-II is robust
From reanalysis of peptide elution studies, we show that 1.4\% of epitopes that are naturally presented on MHC-I 
can be mapped to stem from a TMH. 
We also find that MHC-II presents epitopes derived from TMHs, 
as we find that 3.9-4.0\% of sequenced epitopes
can be uniquely traced onto a human reference proteome (also shown in table \ref{tab:elution}). \geert{zie eerdere vraag. Zijn deze percentages van totale peptides, of van humane peptides?}

% \paragraph{False positives}

A caveat of the detected TMH peptides in elution studies is that we assume that 
the found epitopes are all derived from their completely 
translated predecessor proteins.
However, defective ribosomal 
products (DRiPs) (\cite{yewdell1996defective}),
in which the ribosome terminates its translation prematurely  (\cite{yewdell2011drips}), may result in peptide fragments that did not originate from the mature TMPs.
DRiPs may lead to false positives, as we incorrectly deduce
epitopes to be derived from a TMH, instead of from incompletely formed proteins.

Another caveat is that the TMH prediction tools are imperfect.
This imperfection, however, appears to be of limited importance,
as our data show that two different topology prediction software algorithms give similar results, 
as shown by table \ref{tab:elution}).

A final caveat in the elution studies is that 
there is methodological bias caused by mass spectronomy,
as hydrophobic fragments are detected less efficiently 
by mass spectronomy (\cite{bianchi2017}). 
This means that the number of TMH-derived epitopes
might be underestimated.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{T cells recognize different protein regions than B cells}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% \paragraph{Over-presentation in silico}

An important implication from the over-presentation of TMH-derived epitopes is that T cells will largely recognize 
different protein regions than B cells. 
Presentation of antigens by MHC-II is important 
for the activation of naive B cells by helper T cells. 
For this activation, 
B cells first ingest antigen that is bound to their B cell receptor, 
and subsequently present peptides derived from this antigen in MHC-II 
to helper T cells. 
Following their activation by the T cells, 
B cells mature into plasma cells 
and release antibodies which recognize the same part of the antigen 
as the original B cell receptor. 
B cell receptors and antibodies will thus recognize solvent-exposed regions 
of antigens that are accessible for binding to the B cell receptor. 
However, the results from our study predict that most MHC-II haplotypes 
present relatively hydrophobic peptides,  which are less likely to be solvent-exposed. 
It is unknown why B and T cells seem to predominantly recognize 
different protein regions, but one possibility might be that 
this lowers the chance of B cell mediated autoimmune diseases, 
because auto-reactive B and T cells recognizing different parts 
of the same antigen would need to be present for breakage of B cell tolerance.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Evolutionary conservation of TMHs}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% \paragraph{Selection undetectable in whole proteome}

In general, one might expect that evolutionary selection results in
an immune system that as most attentive for protein regions that are
essential for the survival, proliferation and/or virulence or pathogenic microbes, 
as these will be most conserved.
In SARS-CoV-2, for example, there is preliminary evidence that the strongest
selection pressure is upon residues that change its 
virulence (\cite{velazquez2020positive}).
These regions, however, may only account for a small part of a pathogen's proteome.
Additionally, the structure and function of these essential regions might differ widely between different pathogenic proteins.
Because of this scarcity and variance in targets, 
one can imagine that it will be mostly unfeasible 
to provide innate immune responses against such rare essential protein regions, 
as suggested in a study on influenza (\cite{han2019individual}),
where it was found that the selection pressure
exerted by the immune system was either weak or absent.

% \paragraph{Selection may be detectable in TMHs}

Evolutionary selection of pathogens by a host's immune system,
however, is likelier to occur for proteomic patterns that are general,
over patterns that are rare.
While essential catalytic sites in a pathogenic proteome
might be relatively rare, TMHs are common and thus might be a more feasible 
target for evolution to respond to.
Indeed, we have found the signature of evolution when both factors,
that is, TMHs and catalytic sites are likely to co-occur,
which is in TMPs that span the membrane at least twice.
In contrast to single-spanners, where we found no significant evolutionary conservation, the TMHs of multi-spanners are more evolutionary conserved than soluble protein regions. Likely, the TMHs in many multi-spanners need to interact which each other for correct protein structure and function and they might hence be more structurally constrained compared to the TMHs of single-spanners.
Thus, we speculate that the human immune system is more attentive 
towards TMHs in multi-spanners, as these are evolutionarily more conserved.

There have been more efforts to assess the conservation of TMHs,
using different methodologies.
One such example is \cite{stevens2001substitution}, 
in which aligned protein sequence data was used.
Also this study found that TMHs are evolutionarily more conserved,
as the mean amino acid substitution rate in TMHs is about ten
percent lower,
which is a similar value as we found.
Another example is a study that estimated the conservation
scores for TMHs and soluble regions based on 
alignments of evolutionary related proteins,
and also found that TMHs are more conserved, 
with a conservation score that was 17\% higher in 
TMHs (\cite{oberai2009structural}).
Note that the last study also found that mutations in human TMHs are likelier to cause
a disease, in line with our conclusion that TMHs are more conserved.

% \paragraph{Conclusion}

Together, from this study, two important conclusions can be drawn. 
First, the MHC over-presentation of TMHs is likely a general feature 
and predicted to occur for most haplotypes of both MHC-I and -II 
and for humans as well as bacterial and viral pathogens. 
Second, TMHs are genuinely more evolutionary conserved than soluble protein motifs, 
at least in the human proteome.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Acknowledgments}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

We thank the Center for Information Technology of the University 
of Groningen for its support and for providing access to the Peregrine 
high performance computing cluster. 
FB is funded by a Veni grant from the Netherlands Organization for Scientific
Research (016.Veni.192.026) and an Off-Road Grant from the Dutch Medical Science Foundation (ZonMW 04510011910005).
GvdB is funded by a Young Investigator Grant from 
the Human Frontier Science Program (HFSP; RGY0080/2018), 
and a Vidi grant from 
the Netherlands Organization for Scientific Research (NWO-ALW VIDI 864.14.001). 
GvdB has received funding from the European Research Council (ERC) 
under the European Union’s Horizon 2020 research and 
innovation programme (grant agreement No. 862137.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Data Accessibility}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

All code is archived at \url{http://github.com/richelbilderbeek/someplace},
with DOI \url{https://doi.org/12.3456/zenodo.1234567}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Authors' contributions}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

RJCB and FB conceived the idea for this research. 
MVB helped with the proteome analysis of \emph{M. tuberculosis}.
RJCB wrote the code.
RJCB, FB and GvdB wrote the article.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Bibliography
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% MEE style
\bibliographystyle{mee}
\bibliography{bbbq_article}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\appendix
\section{Supplementary materials}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Differences with Bianchi et al., 2017}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

A part of this study does the same analysis as Bianchi et al., 2017.
mainly concern the use of different
software and a different definition of what an MHC binder is.

% \paragraph{Definition of what a binder is}

The earlier study defined a peptide an MHC binder 
if \emph{within the protein} in which it was found, 
is was among the peptides with the 2\% lowest IC50 values.
This can be seen at \url{https://github.com/richelbilderbeek/bianchi_et_al_2017/blob/master/predict-binders.R},
where the binders are written to file.

However, in this study, an MHC binder is defined as a peptide within a \emph{proteome} in which it is found, that is among the peptides with the 2\% lowest IC50 values.
Subsection \ref{subsec:ic50s_per_haplotype} shows the IC50 values
for a binder per haplotype. We believe that our revised definition is more correct, as it overcomes bias from proteins with very low numbers of peptides and/or MHC-predicted binders.

% \paragraph{Selenoproteins}

Our previous study used the TMHMM web server
to predict TMHs.
The desktop version of TMHMM, however, gives an
error message on the 25 selenoproteins found in the human
reference proteome.
For the sake of reproducible research, we used the desktop version (as
we can call it from scripts) and, due to this, we removed the
selenoproteins from this analysis.

% \paragraph{Compare results in figure}

To verify if the previous and the current method give rise to
notable difference, we show a side-by-side comparison
in figures \ref{fig:bianch_et_al_2017_1a} and \ref{fig:bilderbeek_et_al_2021_1a}.
The figures that haplotypes that over-present or under-present TMH-derived epitopes,
do so in both studies. The extent to which TMH-derived epitopes are
presented, however, is more extreme in our current setup.

\begin{figure}
  \centering
  \begin{subfigure}[t]{0.45\textwidth}
    \centering
    \caption{}
    \includegraphics[width=\linewidth]{bianchi_et_al_2017_results/fig_1a.png}
    \label{fig:bianch_et_al_2017_1a}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.45\textwidth}
    \centering
    \caption{}
    \includegraphics[width=\linewidth]{bbbq_1_smart_results/fig_f_tmh_2_human_mhc1.png}
    \label{fig:bilderbeek_et_al_2021_1a}
  \end{subfigure}

  \caption{
    \textbf{(A)} 
    Results for \cite{bianchi2017}. 
    Red lines denotes the coincidence interval.
    \textbf{(B)}
    Results for this study.
    Red line denotes the percentage as expected by chance.
  }
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Prediction software used}
\label{subsec:prediction_software_used}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{table}[]
  \begin{tabular}{llll}
    Goal & Tool & Reference \\ 
    \hline
    Predict topology                  & TMHMM                     & \cite{krogh2001predicting} \\
    Predict topology                  & PureseqTM                 & \cite{wang2019efficient} \\
    Predict epitopes MHC-I            & \verb;epitope-prediction; & \cite{bianchi2017} \\
    Predict epitopes MHC-II           & NetMHCIIpan               & \cite{nielsen2008quantitative,karosiene2013netmhciipan} \\
    Call TMHMM from R                 & tmhmm                     & \cite{tmhmm} \\
    Call PureseqTM from R             & pureseqtmr                & \cite{pureseqtmr} \\
    Call NetMHCIIpan from R           & netmhc2pan                & \cite{netmhc2pan} \\
    Combine all                       & bbbq                      & \cite{bbbq}
  \end{tabular}
  \caption{
    Overview of all software used in this research.
  }
  \label{table:software_used}
\end{table}

For this research, we needed software to predict protein
topology, as well as the MHC-I and MHC-II binding affinities
of epitopes. We selected our software, by
searching the scientific literature 
to identify the most recent free and open source (FOSS) 
prediction software.
This was done by searching for papers that (1) cite older
prediction software, and (2) present a novel method to make predictions.
As a starting point, per type of prediction software,
a review paper was used (\cite{moller2001evaluation} for protein
topology, \cite{lundegaard2011prediction} for MHC-I
binding affinities and \cite{nielsen2003reliable} for MHC-II binding
affinities).

% \paragraph{TMH prediction}

There are multiple computational tools developed to predict which
parts of a protein forms a TMH.
In 2001, multiple of such prediction tools have been compared (\cite{moller2001evaluation}),
of which TMHMM (\cite{krogh2001predicting}) turned out to be the most accurate, 
as is used in the previous study (\cite{bianchi2017}).
However, TMHMM has a restrictive software license and is nearly two
decades old.
Therefore, PureseqTM (\cite{wang2019efficient}),
was also used in this study, which has been more recently developed
and has a free software license.

% \paragraph{MHC-I epitope prediction}

For MHC-I, there are multiple computational tools developed 
to predict epitopes. 
According to \cite{lundegaard2011prediction}, at that time,
NetMHCcons (\cite{karosiene2012netmhccons}) gave the best predictions.
We used the same tool as used in our earlier study, \verb;epitope-prediction; (\cite{bianchi2017}),

% \paragraph{MHC-II epitope prediction}

Also for MHC-II, there are multiple computational tools developed 
to predict epitopes,
such as using a trained neural network (\cite{nielsen2003reliable})
or a Gibbs sampling approach (\cite{nielsen2004improved}).
According to \cite{lundegaard2011prediction}, in 2011,
from a set of multiple tools, 
NetMHCIIpan (\cite{nielsen2008quantitative,karosiene2013netmhciipan})
made the most accurate predictions.
The most recent FOSS tool available now appears
to be MHCnuggets (\cite{shao2020high}), which can do both MHC-I 
and MHC-II predictions. 
As we already use \verb;epitope-prediction; (\cite{bianchi2017}) 
for MHC-I predictions, we use MHCnuggets only for MHC-II predictions.

% \paragraph{NCBI data retrieval software}

To retrieve the data from the NCBI databases the
\verb;rentrez; R package (\cite{rentrez}) was used
that calls the NCBI website's API. To provide for a 
stable user experience for all users, 
this API limits the user to 3 calls per second.
Additionally, the API splits the result of a bigger
query into multiple pages, each of which needs one API call.
We wrote the \verb;sprentrez; package (\cite{sprentrez}) to provide for 
bigger queries of multiple (and delayed) API calls.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Prediction software written}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The R programming language is used for the complete 
experiment, including the analysis.
The complete experiment is bundled in the 'bbbq' R package,
which is dependent on 'tmhmm', 'pureseqtmr', 
'epitope-prediction' and 'mhcnuggetsr'
as described below.

% \paragraph{tmhmm}

The R package 'tmhmm' was developed to do the similar topology
predictions as our earlier study (that used 'TMHMM'), yet in an automated way.
'TMHMM' has a restrictive software license (\cite{krogh2001predicting}) and allows a user
to download a pre-compiled executable after confirmation he/she
is in academia. The R package respects this restriction
and allows the user to install and use TMHMM from within R,
as done in this study.
'tmhmm' has been submitted to and is accepted 
by the Comprehensive R Archive Network (CRAN).

% \paragraph{pureseqtmr}

To be able to call, from R, the TMH prediction 
software 'PureseqTM' (\cite{wang2019efficient}),
which is written in C, the package 'pureseqtmr' has been developed. 
'pureseqtmr' allows to install 'PureseqTM' and use most of its features.
'pureseqtmr' has been submitted to and is accepted by CRAN.

% \paragraph{mhcnuggetsr}

MHCnuggets is a free and open-source Python package to predict 
epitope affinity for many MHC-I and MHC-II variants (\cite{shao2020high}).
The R package 'mhcnuggetsr' allows one to install and use MHCnuggets
from within R.
Also 'mhcnuggetsr' has been submitted to and is accepted by CRAN.

% \paragraph{bbbq}

To reproduce the full experiment presented in this paper,
the functions needed are bundled in the 'bbbq' R package.
This package is too specific to be submitted to CRAN.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Prediction of percentage of epitopes overlapping with a TMH}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Table \ref{tab:f_tmh} shows an overview of the findings,
where a target specifies the source of the proteome,
where \verb;covid; denotes SARS-CoV-2 and \verb;myco; denotes
Mycobacterium tuberculosis. \verb;mhc_class; denotes the MHC
class, \verb;n_spots; the number of possible 9-mers (for MHC-I) 
or 14-mers (for MHC-II) possible. \verb;n_spots_tmh; the
number of epitopes that overlapped with a TMH that were binders. 
\verb;f_tmh; the percentage of peptides that had at least 1 residue
overlapping with a TMH.

% Label: tab:f_tmh
\input{bbbq_1_smart_results/table_f_tmh_2.latex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Minor methods}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

These are details that are removed from the 'Methods' section.

PureseqTM does not predict the topology
of proteins that have less than three amino acids. 
The TRDD1 ('T cell receptor delta diversity 1') protein,
however, is two amino acids long. 
The R package \verb;pureseqtmr;, however, 
predicts that mono- and di-peptides are cytosolic.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Minor discussion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

These are details that are removed from the 'Discussion' section.

% \paragraph{Bacteria have a different cell membrane}

In this experiment we predicted epitopes that overlap with 
TMHs from a human, bacterial and viral proteome,
would these proteins be expressed in a human host.
Bacteria, however have different cell membranes and cell walls, 
hence different structural requirements for a TMH.
Both topology prediction tools were trained to recognize
human TMHs, thus we cannot be sure that
the transmembrane regions predicted in bacterial proteins
are actually part of a TMH.
For the purpose of this study, we assume the 
error in topology predictions to be unbiased way towards topology.
In other words: that a bacterial TMH is incorrectly
predicted to be absent just as often as it is incorrectly
predicted to be present elsewhere.

% \paragraph{False positives in SNPs}

Regarding the evolutionary conservation of TMHs using SNPs,
again, it is estimated that approximately ten percent
of SNPs is a false positive that result from the methods to determine
a SNP. One example is that sequence variations are incorrectly
detected due to highly similar duplicated sequences \cite{musumeci2010single}.
We assume that these duplications occur as often in TMHs as in
regions around these, hence we expect this not to affect our results.

% \paragraph{synonymous mutations}
%
In our evolutionary experiment, 
we removed variations that were synonymous mutations (i.e.
resulted in the same amino acid, from a different genetic code) 
from our analysis.
There is evidence, however, that these synonymous mutations
do have an effect and may even be evolutionary selected 
for (\cite{hunt2009silent}).
As the possible effect of synonymous mutations is ignored by our
topology prediction software, we do so as well.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Elution studies}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% tab:elution
\input{bbbq_article_issue_157/results.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{IC50 values of binders per haplotype}
\label{subsec:ic50s_per_haplotype}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Per target proteome (i.e. human, SARS-CoV-2, Mycobacterium tuberculosis),
we collected all 9-mers (for MHC-I) and 14-mers (for MHC-II),
after removing the selenoproteins and proteins that are shorter
than the epitope length.
From these epitopes, per MHC haplotype,
we predicted the IC50 (in nM) using \verb;epitope-prediction; (for MHC-I)
and MHCnuggets (for MHC-II). 
Here, we show the IC50 value per haplotype that
is used to determine if a peptide binds to the haplotype's MHC
for MHC-I (see table \ref{tab:ic50_binders_mhc1}) and 
MHC-II (see table \ref{tab:ic50_binders_mhc2}).

% tab:ic50_binders_mhc1
\input{bbbq_1_smart_results/table_ic50_binders_mhc1_2.latex}

% tab:ic50_binders_mhc2
\input{bbbq_1_smart_results/table_ic50_binders_mhc2_2.latex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Presentation of TMH-derived epitopes}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Table \ref{tab:haplotype_abbreviations} shows the
shorthand notation for the HLA haplotypes.

% Label: tab:haplotype_abbreviations
\input{bbbq_1_smart_results/haplotype_abbreviations.tex}

Tables \ref{tab:tmh_binders_mhc1} and \ref{tab:tmh_binders_mhc2}
show the exact number of binders, binders that overlap with TMHs
and the percentage of binders that overlap with TMHs, as
visualized by figure \ref{fig:bbbq_1_smart_results}.

% Label: tab:tmh_binders_mhc1
\input{bbbq_1_smart_results/table_tmh_binders_mhc1_2.latex}

% Label: tab:tmh_binders_mhc2
\input{bbbq_1_smart_results/table_tmh_binders_mhc2_2.latex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Relative presentation of TMH-derived epitopes}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

To compare the over-presentation of TMH-derived epitopes between the
different proteomes, we normalized this percentages in such a
way that 1.0 is the percentage of TMH-derived epitopes that would 
be expected by chance. 
Figure \ref{fig:f_tmh_mhc1_normalized} and \ref{fig:f_tmh_mhc2_normalized}
show these normalized values for the MHC-I and MHC-II haplotypes respectively.

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{bbbq_1_smart_results/fig_f_tmh_mhc1_2_normalized.png}
  \caption{
    Normalized proportion of MHC-I epitopes overlapping with TMHs
    for human, viral and bacterial proteomes.
    Legend: covid = SARS-CoV-2,
    human = homo sapiens, myco = Mycobacterium tuberculosis
  }
  \label{fig:f_tmh_mhc1_normalized}
\end{figure}

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{bbbq_1_smart_results/fig_f_tmh_mhc2_2_normalized.png}
  \caption{
    Normalized proportion of MHC-II epitopes overlapping with TMHs
    for human, viral and bacterial proteomes.
    Legend: covid = SARS-CoV-2,
    human = homo sapiens, myco = Mycobacterium tuberculosis
  }
  \label{fig:f_tmh_mhc2_normalized}
\end{figure}

To determine the additional over-presentation of TMH-derived epitopes 
in MHC-II (as compared to MHC-I), we normalized the data to enable
a side-by-side comparison. 
The percentage of TMH-derived epitopes presented was normalized
to the expected percentage of TMH-derived epitopes,
where $1.0$ denotes that the percentage of presented TMH-derived epitopes
matches the values as expected by chance.
The normalized values per haplotype are shown 
in figure \ref{fig:rel_presentation_per_haplotype}.
To compare the TMH-derived over-presentation per MHC class,
we grouped the normalized values per haplotype, 
and plot the mean and standard error, as shown in figure \ref{fig:rel_presentation}.

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{bbbq_1_smart_results/fig_rel_presentation_per_haplotype.png}
  \caption{
    Normalized proportion of MHC-I and MHC-II epitopes overlapping with TMHs,
    for the different haplotypes and proteomes
  }
  \label{fig:rel_presentation_per_haplotype}
\end{figure}

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{bbbq_1_smart_results/fig_rel_presentation.png}
  \caption{
    Normalized proportion of MHC-I and MHC-II epitopes overlapping with TMHs,
    for the different MHC classes and proteomes. Error bars denote the
    standard error.
  }
  \label{fig:rel_presentation}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Evolutionary conservation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

See tables \ref{tab:ncbi_counts_1} and tables \ref{tab:ncbi_counts_2}
for an overview of all amounts.
In table \ref{tab:ncbi_counts_1} there are multiple instances where
the amounts are expected to add up, yet don't, as one SNP can work on
multiple isoforms. For example, there are 9,621 unique SNPs 
found in all proteins, of which 4,219 around found in MAPs 
and 6,026 in TMPs. Apparently, 624 SNPs work on a set of isoforms that
contains both MAPs and TMPs.

% Label: tab:ncbi_counts_1
\input{ncbi_peregrine_results/ncbi_counts_1.tex}

% Label: tab:ncbi_counts_2
\input{ncbi_peregrine_results/ncbi_counts_2.tex}

Figure \ref{fig:snps_per_gene_name_ncbi} shows the distribution of the
number of SNPs per gene name, at 2020-12-14.

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_snps_per_gene_name_ncbi.png}
  \caption{
    Distribution of the number of SNPs per gene name in the NCBI database.
  }
  \label{fig:snps_per_gene_name_ncbi}
\end{figure}

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_snps_per_gene_name_processed.png}
  \caption{
    Distribution of the number of protein variations and SNPs per gene name processed.
  }
  \label{fig:snps_per_gene_name_processed}
\end{figure}

%Figure \ref{fig:f_tmh_ncbi} shows the distribution 
%of the percentages of TMH of transmembrane proteins.
%
%\begin{figure}[!htbp]
%  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_f_tmh_ncbi.png}
%  \caption{
%    Distribution of the percentages of TMH of transmembrane proteins.
%    Dashed vertical line: mean \% TMH in all proteins.
%    Dotted vertical line: mean \% TMH in TMPs.
%  }
%  \label{fig:f_tmh_ncbi}
%\end{figure}

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_n_proteins_per_gene_name.png}
  \caption{
    Histogram of the number of proteins found per gene name.
    Most often, a gene name is associated with one proteins. 
  }
  \label{fig:n_proteins_per_gene_name}
\end{figure}

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_n_snps_per_tmp.png}
  \caption{
    Histogram of the number of SNPs per trans-membrane protein.
    Dashed vertical line: average number of SNPs per TMP
  }
  \label{fig:fig_n_snps_per_tmp}
\end{figure}

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_n_tmhs_per_protein.png}
  \caption{
    Histogram of the number of TMHs predicted per protein,
    for the trans-membrane proteins used.
  }
  \label{fig:fig_n_tmhs_per_protein}
\end{figure}

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_snp_rel_pos.png}
  \caption{
    Distribution of the relative position of the SNPs used,
    where a relative position of zero denotes the first amino
    acid at the N-terminus, where a relative position of one
    indicates the last residue at the C-terminus.
  }
  \label{fig:snp_rel_pos}
\end{figure}

% \paragraph{Relative position of SNPs}

To verify if SNPs were sampled uniformly
over proteins, we show the distribution 
of the relative position in figure \ref{fig:snp_rel_pos}.
We find no clear evidence of a bias.

% \paragraph{Statistics}

Table \ref{tab:snp_stats} shows the statistics for all
SNPs, where tables \ref{tab:snp_stats_per_spanner_single}
and \ref{tab:snp_stats_per_spanner_multi} show the
statistics for only single-spanners and multi-spanners respectively.

% Label: tab:snp_stats
\input{ncbi_peregrine_results/snp_stats.tex}

% Label: tab:snp_stats_per_spanner_single
\input{ncbi_peregrine_results/snp_stats_per_spanner_single.tex}

% Label: tab:snp_stats_per_spanner_multi
\input{ncbi_peregrine_results/snp_stats_per_spanner_multi.tex}

%
% Notes to self
%

%\begin{sidewaystable}
%  \centering
%  ... centered table here
%\end{sidewaystable}

% \scalebox{0.7}{
%  ... scaled table here
% }

%{\tiny
%  ... something with a tiny font
%}

% Instead of non-TMH use 'soluble protein region'
% or 'cytosolic or extracellular protein region'