### 2021-10-29 bbbq_article_issue_157/results.tex
Add two immunopeptidomes studies, progress #229

% latex table generated in R 4.1.1 by xtable 1.8-4 package
% Fri Oct 29 12:40:25 2021
\begin{table}[ht]
\centering
\begin{tabular}{llll}
  \hline
MHC class & Tool & Dataset & n \\ 
  \hline
I & PureseqTM & schellens & 1.38\% (109/7897) \\ 
  I & PureseqTM & iedb & 6.81\% (43/631) \\ 
  I & TMHMM & schellens & 1.43\% (113/7897) \\ 
  I & TMHMM & iedb & 7.13\% (45/631) \\ 
  II & PureseqTM & bergseng & 3.92\% (498/12712) \\ 
  II & PureseqTM & iedb & 0.29\% (4/1364) \\ 
  II & TMHMM & bergseng & 3.96\% (504/12712) \\ 
  II & TMHMM & iedb & 1.39\% (19/1364) \\ 
   \hline
\end{tabular}
\caption{Percentage of epitopes derived from a TMH found in the two elution studies, for the two different kind of topology prediction tools. The values between braces show the the number of epitopes that were predicted to overlapping with a TMH per all epitopes that could be uniquely mapped to the representative human reference proteome.} 
\label{tab:elution}
\end{table}